---
title: "Assignment 1 Task 1"
author: "Sarah Lam"
date: "1/14/2022"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readxl)
library(here)
library(lubridate)
library(patchwork)
library(cowplot)
```

## Overview

Data summary: You will explore amphibian abundance data recorded by the Sierra Lakes Inventory Project. From the Environmental Data Initiative repository: “The Sierra Lakes Inventory Project (SLIP) was a research endeavor that ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. We described the physical characteristics of and surveyed aquatic communities for >8,000 lentic water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows.”

Data citation: Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

```{r} 
#read in data
amphibians <- read_excel(here("data", "sierra_amphibians.xlsx"), col_types = "text")
```

```{r, fig.cap = "The figure shows the onserved counts of Mountain Yellow-Legged Frogs each years from 1995 to 2002 for each stage of the frog life cycle, tapole, subadult, and adult. All three lines follow a similar trend with an overall increase in observed counts after a decrease in 2001 and adult frogs (yellow) having the highest observed numbers."}
#data wrangling
yellow_leg_frog <- amphibians %>% 
  filter(amphibian_species == "RAMU") %>% # filter for just the frog species we want
  filter(amphibian_life_stage != "EggMass") %>% #remove egg mass life stage since we won't graph it 
  mutate(date_class = ymd(survey_date)) %>% #change class of survey date
  mutate(year = year(date_class)) %>%  #pull out year from date
  group_by(year, amphibian_life_stage) %>% #group by year and life stage
  summarize(count = n()) #count frogs recorded at each life stage each year

# plot data
observe_overtime <- ggplot(data = yellow_leg_frog, aes(x = year, y = count)) +
  geom_line(aes(color = amphibian_life_stage)) +
  scale_color_manual(values = c("goldenrod2", "coral2", "cyan3")) +
  labs(x = "Year", 
       y = "Number of Frogs Observed", 
       title = "Mountain Yellow-Legged Frog Observations, 1995-2002 (Lam)", 
       color = "Frog Life Stage") +
  theme_minimal()
```

```{r, fig.cap = "The figure shows, in descending order, the 5 lakes with the highest counts of observed adult and sub adult mountain yellow-legged frogs."}
# total counts by lake 
yellow_frogs_by_lake <- amphibians %>% 
  filter(amphibian_species == "RAMU", 
         amphibian_life_stage %in% c("SubAdult", "Adult")) %>% 
  group_by(lake_id) %>% 
  summarize(count = n()) %>% 
  slice_max(count, n = 5) %>% 
  mutate(lake_id = as.factor(lake_id)) %>% 
  mutate(lake_id = fct_reorder(lake_id, desc(count)))
  
count_bylake <- ggplot(data = yellow_frogs_by_lake, aes(x = lake_id, y = count)) +
  geom_col(fill = "gold3") +
  labs(x = "Lake Identification Label",
       y = "Number of Frogs Observed",
       title = "Count of Adult and Sub Adult Mountain Yellow-Legged Frogs in Top 5 Lakes (Lam)") +
  theme_minimal()
```

```{r}
observe_overtime / 
  count_bylake
```

A column graph containing total counts (over all years of the study) of combined adult and subadult endangered mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed counts. In other words, this graph will have at least 5 columns (OK to have more if there are ties), with Lake ID (these aren’t specified in the dataset for confidentiality) as the categorical label on one axis, and total yellow-legged frog counts (adult + subadult counts) in the dataset on the other axis. Make sure they’re in high-to-low or low-to-high order by total yellow-legged frog counts. You should exclude tadpoles for this graph. Note: Consider reformatting the lake ID from just a number to a label like “Lake 10025”), then use fct_reorder to convert it to an ordered factor.

Combine your two graphs into a single compound figure using the {patchwork} or {cowplot} package. Add a finalized figure caption that appears below the compound figure in your knitted html. Note: You might realize once you combine your graphs into a single compound figure, you need to update the formatting (e.g. move legends or instead directly label, etc.). You should customize as necessary to make it a professional final output.

For Task 1, your knitted html should show: 
Your organized code, with clear subsections and any useful descriptive text / annotation (e.g. if you wanted to highlight this as a code example for a prospective employer). Include an overview of the data and your exploration, and a citation for the data.
Your finalized compound figure, with a figure caption.
Make sure to suppress any messages & warnings, but include the setup chunk to show which packages you’ve loaded.

Submit your Task 1 knitted html on GauchoSpace: 
a1_task1_lastname_firstname.html
