---
title: "Assignment 1 Task 1"
author: "Sarah Lam"
output: html_document
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(readxl)
library(here)
library(lubridate)
library(patchwork)
library(cowplot)
```

## Overview

This code explores amphibian abundance data recorded by the Sierra Lakes Inventory Project (SLIP) from 1995-2000. This code focuses on Rana Muscosa, or mountain yellow-legged frog, abundance over of the five year time frame and across the major amphibian life stages. For this purposes of this data visualization, both dead and alive observed specimens have been accounted for under the assumption that dead specimens were recently deceased and because the dead observations were a small proportion of observed individuals. 

Data citation: Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3

```{r} 
#read in data
amphibians <- read_excel(here("data", "sierra_amphibians.xlsx"), col_types = "text")
```

```{r}
#data wrangling
yellow_leg_frog <- amphibians %>% 
  filter(amphibian_species == "RAMU") %>% # filter for just the frog species we want
  filter(amphibian_life_stage != "EggMass") %>% #remove egg mass life stage since we won't graph it 
  mutate(date_class = ymd(survey_date)) %>% #change class of survey date
  mutate(year = year(date_class)) %>%  #pull out year from date 
  mutate(amphibian_number = suppressWarnings(as.numeric(amphibian_number))) %>% 
  group_by(year, amphibian_life_stage) %>% #group by year and life stage
  summarize(count = sum(amphibian_number, na.rm = TRUE)) #count frogs recorded at each life stage each year

# plot data
observe_overtime <- ggplot(data = yellow_leg_frog, aes(x = year, y = count)) +
  geom_line(aes(color = amphibian_life_stage)) +
  scale_color_manual(values = c("goldenrod2", "coral2", "cyan3")) +
  labs(x = "Year", 
       y = "Number of Frogs Observed", 
       title = "Mountain Yellow-Legged Frog Observations, 1995-2002 (Lam)", 
       color = "Frog Life Stage") +
  theme_minimal()
```

```{r}
# total counts by lake 
yellow_frogs_by_lake <- amphibians %>% 
  filter(amphibian_species == "RAMU", 
         amphibian_life_stage %in% c("SubAdult", "Adult")) %>% 
  select(lake_id, amphibian_number) %>% 
  mutate(amphibian_number = suppressWarnings(as.numeric(amphibian_number))) %>% 
  group_by(lake_id) %>% 
  summarize(total_count = sum(amphibian_number, na.rm = TRUE))  %>% 
  slice_max(total_count, n = 5) %>% 
  mutate(lake_id = paste('Lake', lake_id, sep = ' ', collapse = NULL)) %>% 
  mutate(lake_id = as.factor(lake_id)) %>% 
  mutate(lake_id = fct_reorder(lake_id, desc(total_count))) 
  
 
  
count_bylake <- ggplot(data = yellow_frogs_by_lake, aes(x = lake_id, y = total_count)) +
  geom_col(fill = "gold3") +
  labs(x = "Lake Identification Label",
       y = "Number of Frogs Observed",
       title = "Count of Adult and Sub Adult Mountain Yellow-Legged Frogs in Top 5 Lakes (Lam)") +
  theme_minimal()
```

```{r}
observe_overtime / 
  count_bylake
```

**Fig. 1.** Two graphs visualizing mountain yellow-legged frog populations in the Sierra Nevada region of California. The first figure (above) shows the observed counts of Mountain Yellow-Legged Frogs each years from 1995 to 2002 for each stage of the frog life cycle, tadpole, sub adult, and adult. All three lines follow a similar trend, however tadpoles(blue) clearly show the highest observed counts and the greatest variations in numbers overtime.
The second figure (below) shows, in descending order, the 5 lakes with the highest counts of observed adult and sub adult mountain yellow-legged frogs, lake names are omitted here for confidentiality reasons.
